### 2022-10-01 Papers/ICCC-2023/sections/design.tex
stuff
.

@@

CTFE First apporach was created during implementation of the first compiler for the C-=-1 language\cite{grabski2022compilation}.
CTFE First compiler has four major components: Frontend, Interpreter, Compiler Interface and Backend.
Figure \ref{CTFE-first-compiler-structure} contains a diagram with an overview of how these pieces interact with eachother, during the compilaton process.
Frontend, described in chapter \ref{frontend}, parses the code in the compiled language and constructs its intermidaite representation, using interpreter's data structures.
It is used to analyze both user code and the Compiler Interface.
After the intermidiate representation is constructed, it is passed on to the Interpreter, which was described in chapter \ref{interpreter}.
Compiler Interface intermidiate representation is then executed, using the user program as data. 
This step converts the semantic model of the program into the backends intermidiate language.
This process has been further explained in chapter \ref{compiler-interface}.
Finaly the Backend generates the executable file.

@@

@@

Compiler Interface translates the program's intermidaite representation into the backend's assembly language.
This component is interpreted during compilation.
What stands out about CTFE First is that this part of the compiler can thus be written in the target language, during Stage 0 of the compiler bootstrapping process, as was the case for the C-=-1 compiler\cite{puntambekar:compiler_design, novillo2007gcc, grabski2022compilation}.

@@

CTFE First does not put any additional requierments on compiler backend.
When using this pattern, an out-of-the-box backend library, can be used, as was the case with C-=-1 compiler.

The backend code must be invokable from within the interpreted program in the target language.
Depending on how the interpreter was designed, this can represent a significant undertaking.
Compiler backends are large and for the Compiler Interface to take advantage of them fully, they must be fully available.
This means exposing each function and type within the library, to the interpreted code.
These bindings could feasibly be generated automatically \cite{marshalling_auto_generation}, but this tehnique was not used when implementing C-=-1 compiler.

>>>>

C-=-1 type system heavily borrows from C++.

@@

CTFE First apporach was created during implementation of the first compiler for the C-=-1 language\cite{grabski2022compilation}.
CTFE First compiler has four major components: Frontend, Interpreter, Compiler Interface and Backend.
Figure \ref{CTFE-first-compiler-structure} contains a diagram with an overview of how these parts interact with eachother, during the compilaton process.
Frontend, described in chapter \ref{frontend}, parses the code in the compiled language and constructs its intermidaite representation, using interpreter's data structures.
It is used to analyze both user code and the Compiler Interface.
After the intermidiate representation is constructed, it is passed on to the Interpreter, which was described in chapter \ref{interpreter}.
Compiler Interface intermidiate representation is then executed, using the user program as data.
This step converts the semantic model of the program into the backends intermidiate language.
This process is further explained in chapter \ref{compiler-interface}.
Finaly the Backend generates the executable file.

@@

An important consideration with implementing a CTFE First compiler, is what limitations to put on modifications of user code.
Depending on language design, it may be possible to introduce circular references between the functions that modify the codebase.
For example if function \lstinline{A} is modified by function \lstinline{B} and \lstinline{B} invokes \lstinline{A}, the behaviour of the program is unpredictable.
This problem will only be magnified by larger program sizes.

One of the possible solutions to the above mentioned problem is to limit which functions can be invoked at compiletime.
C-=-1 allows code within a compiletime context to invoke procedures only from other packages, declared explicitly as dependencies.
It additionally prohibits modification of dependencies.
Therefore its is impossible for a function to modify a procedure it depends on.% todo: wording, example

@@

Compiler Interface translates the program's intermidaite representation into the backend's assembly language.
This component is interpreted during compilation.
What is unique about CTFE First is that this part of the compiler can thus be written in the target language, during Stage 0 of the compiler bootstrapping process, as was the case for the C-=-1 compiler\cite{puntambekar:compiler_design, novillo2007gcc, grabski2022compilation}.

@@

CTFE First does not put any additional requirements on compiler backend.
When using this pattern, an out-of-the-box backend library, can be used, as was the case with C-=-1 compiler.

The backend code must be invokable from within the interpreted program in the target language.
Depending on how the interpreter was designed, this can represent a significant undertaking.
Compiler backends are large and for the Compiler Interface to take advantage of them, they must be fully available.
This means exposing each function and type within the library, to the interpreted code.
These bindings could feasibly be generated automatically \cite{marshalling_auto_generation}, but this tehnique was not used when implementing C-=-1 compiler.

### 2022-10-01 Papers/ICCC-2023/sections/realted-work.tex
.

Rust macros, mutch like the classic C-style preprocessor macros, generate additional code as text.
The major difference between the systems is that Rust macros operate on tokens rather than text and use syntax similar to regular Rust code.
This mechanism is powerfull, allowing the user to rewrite code to avoid repetition, simplify certain tasks and create new syntax.
They are however unable to reflect on the program as a whole or obtain mutch of the information available to the compiler.

Build scripts, on the other hand, are programs that execute prior to compilation of the main package.
They are meant to prepare the envoirment for building the program, by for example compiling its external dependencies.
They do not have any information about the structure of the program being compiled.
If a task, done by a build script, were to require such data, it would have to analyze the code by itself, without any help from the compiler.
This makes application such as automatically genertaing bindings for other programming languages significantly more difficult.

>>>>

Rust macros, much like the classic C-style preprocessor macros, generate additional code as text.
The major difference between the systems is that Rust macros operate on tokens rather than text and use syntax similar to regular Rust code.
This mechanism is powerfull, allowing the user to rewrite code to avoid repetition, simplify certain tasks and create new syntax.
They are however unable to reflect on the program as a whole or obtain much of the information available to the compiler.

Build scripts, on the other hand, are programs that execute prior to compilation of the main package.
They preparing the envoirment for building the program by, for example, compiling its external dependencies.
They do not have any information about the structure of the program being compiled.
If a task, done by a build script, were to require such data, it would have to analyze the code by itself, without any help from the compiler.
This makes application such as automatically genertaing bindings for other programming languages significantly more difficult.
